---
title: "QTM350 - Homework 5: World Development Indicators Analysis"
author: "Yunshi Zhou"
format:
  revealjs:
    theme: moon
    transition: slide
    slide-number: true
---

## Introduction

This presentation explores selected indicators from the World Development Indicators dataset.

Key indicators analysed:

-   GDP per capita
-   Life expectancy
-   Primary school enrollment rate

## Income and Health Relationship

```{r}
#| echo: false
library(tidyverse)
wdi <- read_csv("wdi.csv")
library(ggplot2)
ggplot(wdi, aes(x = gdp_per_capita, y = life_expectancy)) +
  geom_point(alpha = 0.6) +
  labs(
    title = "GDP per Capita vs Life Expectancy",
    x = "GDP per Capita",
    y = "Life Expectancy"
  )
```

## Primary School Enrollment

```{r}
#| echo: false
top10 <- wdi %>%
  arrange(desc(primary_school_enrolment_rate)) %>%
  slice(1:10)
ggplot(top10, aes(x = reorder(country, primary_school_enrolment_rate), y = primary_school_enrolment_rate)) +
  geom_col() +
  coord_flip()
```

## Key Findings

-   Higher GDP is associated with longer life expectancy.
-   Enrollment rates are generally high worldwide.
-   Economic inequality remains significant.

## Conclusion

This analysis highlights global disparities in income, health, and education.

Future work could explore trends over time and regional patterns.

Thank you!
